Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-18 2017-Xudong-SMT/FM-2018-extended.tex
Additional example of priority in Section 3 Moved the figures a bit in Section 5

As a running example we use the Failure Monitor architecture from the
CubETH nanosatellite on-board software
case-study~\cite{CubETH-case-study} realised using BIP.  BIP
(Behaviour-Interaction-Priority)~\cite{bip} is a framework for the
component-based design of concurrent software and systems.  In
particular, the BIP tool-set comprises compilers for generating C/C++
code, executable by linking with one of the dedicated engines, which
implement the BIP operational semantics~\cite{BliSif08-acp-tc}.  This
approach ensures that any property, shown to hold on a given BIP
model, will also hold by construction on the generated code.

@@

Finally, \emph{priorities}\mdash defined by a strict partial order on
the set of possible interactions\mdash narrow the choice among the
enabled interactions at any given execution round.  The default priority is the
so-called \emph{maximal progress}, whereby among any two interactions
$a \subset b$ (as sets of ports), $b$ has higher priority than $a$.
For example, the port $\NameOprnd.\PortFail$ will never fire alone in
a global state, where both $\NameTimer.\PortStart$ and
$\NameCtrl.\PortFail$ are enabled.

Application of the Failure Monitor architecture ensures that, whenever
a failure is registered in the operand component, the system will be
reset, unless a resumption is registered within $\mathrm{Max}$ time
units.

@@

Figure~\ref{schema:ArchFailure:pNet} shows a pNet encoding of the
above Failure Monitor architecture.  This encoding is structural: each
coordinator component is encoded as a pLTS, the operand
component\mdash as a hole; connectors of the BIP model are encoded as
synchronisation vectors.  Each connector that does not involve
triggers is trivially encoded by a synchronisation vector comprising
the same ports.  In order to encode the semantics of the connectors
involving triggers, we 1)~in the pLTS encoding the coordinator
components, add loop transitions to ensure that all ports involved in
such connectors are enabled in all states, 2)~associate a Boolean
value to each of these ports: the original transitions carry the value
$\true$ (\eg $\cstate{0} \goesto{\PortFail(\true)} \cstate{1}$), the
added loops carry the value $\false$ (\eg $\cstate{2}
\goesto{\PortFail(\false)} \cstate{2}$), 3)~add to the corresponding
synchronisation vectors the Boolean predicate encoding the connector
structure.  For example, $SV0$ encodes the connector discussed above:
the predicate $(b1=b2) \land (b1\lor b2 \Rightarrow b0)$ means that the
``true'' transitions $\NameCtrl.\PortFail$ and $\NameTimer.\PortStart$
can only fire together ($b1 = b2$) and whenever one of them fires,
$\NameOprnd.\PortFail$ must fire also ($b1 \lor b2 \Rightarrow b0$).
This encoding can be systematically obtained for any hierarchical BIP
connector~\cite{BliSif10-causal-fmsd}.
%
Although, for the sake of brevity, this encoding does not translate
the maximal progress priority, it can be easily extended to do so, by
introducing an additional Boolean variable for each relevant port.

@@

\begin{figure}[h]
\begin{mathpar}
    \openrule{\{s1 \xrightarrow{\OTvar{fail(false)}} s1, t1 \xrightarrow{\OTvar{resume(true)}} t0\}, ~~\{\xrightarrow{hb}\},~~\OTvar{fail(false)}=\OTvar{fail(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{resume(true)}=\OTvar{start(}\OTvar{b}_2\OTvar{)}\OTland \OTvar{hb}=\OTvar{fail(}\OTvar{b}_0\OTvar{)}\OTland \OTvar{v}=\OTvar{\underline{fail}}\OTland \OTvar{b}_1=\OTvar{b}_2\\
    \OTland \OTvar{(}\OTvar{b}_1\lor\OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{\}}
    	   {\ostate{s1,t1} \xrightarrow{\OTvar{v}} \ostate{s1,t0}}\\
  \end{mathpar}
\caption{One of the unsatisfiable open transitions in FailureTimer}  \label{schema:unsat-ot}
\end{figure}

@@

@@

Fig.~\ref{schema:smt-lib} shows the translation of the
transition of Fig.~\ref{schema:unsat-ot}
in the SMTlib syntax.
It contains the declaration
of the BIP action algebra sorts 
and constructors, then the declaration of variables, and finally the
predicate to be checked, encoded as a set of assertions.
Here we display also the proof generated by Z3,
though in practice our tool only needs the brute information ``sat''
or ``unsat''. The result "unsat" in the output is just what we expected.
% as appears in the ``unsat'' proof generated by Z3 in which ``\{= ?x66 ?x70\}'' is shown to be false (see Fig. \ref{schema:smt-lib})

@@

\subsection{Result of the running example}
\label{section:resultOA}

@@

\begin{figure}[t]
   \centerline{\includegraphics[width=12cm]{XFIG/FailureTimerOA}}
   \caption{Open Automaton}  \label{schema:resultOA}
 \end{figure}

>>>>

As a running example we use the Failure Monitor architecture from the
CubETH nanosatellite on-board software
case-study~\cite{CubETH-case-study} realised using BIP.  BIP
(Behaviour-Interaction-Priority)~\cite{bip} is a framework for the
component-based design of concurrent software and systems.  In
particular, the BIP tool-set comprises compilers for generating C/C++
code, executable by linking with one of the dedicated engines, which
implement the BIP operational semantics~\cite{BarBliu15-offer-scico}.  This
approach ensures that any property, shown to hold on a given BIP
model, will also hold by construction on the generated code.

@@

Finally, \emph{priorities}\mdash defined by a strict partial order on
the set of possible interactions\mdash narrow the choice among the
enabled interactions at any given round.  The default priority is the
so-called \emph{maximal progress}, whereby among any two interactions
$a \subset b$ (as sets of ports), $b$ has higher priority than $a$.
For example, the port $\NameOprnd.\PortFail$ will never fire alone in
a global state, where both $\NameTimer.\PortStart$ and
$\NameCtrl.\PortFail$ are enabled.
%
Adding priorities $\NameOprnd.\PortFinish < \NameCtrl.\PortReset$ and
$\NameOprnd.\PortFinish < \NameOprnd.\PortResume$, ensures that, after
a failure, the operation performed by $\NameOprnd$ cannot finish
unless it is explicitly resumed or the system is reset.

Application of the Failure Monitor architecture ensures that, whenever
a failure is registered in the operand component, the system will be
reset, unless a resumption is registered within $\mathrm{Max}$ time
units (more details in Sect.~\ref{section:resultOA}).

@@

Figure~\ref{schema:ArchFailure:pNet} shows a pNet encoding of the
above Failure Monitor architecture.  This encoding is structural: each
coordinator component is encoded as a pLTS, the operand
component\mdash as a hole; connectors of the BIP model are encoded as
synchronisation vectors.  Each connector that does not involve
triggers is trivially encoded by a synchronisation vector comprising
the same ports.  In order to encode the semantics of the connectors
involving triggers, we 1)~in the pLTS encoding the coordinator
components, add loop transitions to ensure that all ports involved in
such connectors are enabled in all states, 2)~associate a Boolean
value to each of these ports: the original transitions carry the value
$\true$ (\eg $\cstate{0} \goesto{\PortFail(\true)} \cstate{1}$), the
added loops carry the value $\false$ (\eg $\cstate{2}
\goesto{\PortFail(\false)} \cstate{2}$), 3)~add to the corresponding
synchronisation vectors the Boolean predicate encoding the connector
structure.  For example, $SV0$ encodes the connector discussed above:
the predicate $(b1=b2) \land (b1\lor b2 \Rightarrow b0)$ means that the
``true'' transitions $\NameCtrl.\PortFail$ and $\NameTimer.\PortStart$
can only fire together ($b1 = b2$) and whenever one of them fires,
$\NameOprnd.\PortFail$ must fire also ($b1 \lor b2 \Rightarrow b0$).
This encoding can be systematically obtained for any hierarchical BIP
connector~\cite{BliSif10-causal-fmsd}.
%
Although, for the sake of brevity, we omit priorities from the
encoding, this can be easily achieved, by introducing additional
Boolean variables for relevant ports~\cite{BarBliu15-offer-scico}.

@@

@@

\begin{figure}[t]
\begin{mathpar}
    \openrule{\{s1 \xrightarrow{\OTvar{fail(false)}} s1, t1 \xrightarrow{\OTvar{resume(true)}} t0\}, ~~\{\xrightarrow{hb}\},~~\OTvar{fail(false)}=\OTvar{fail(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{resume(true)}=\OTvar{start(}\OTvar{b}_2\OTvar{)}\OTland \OTvar{hb}=\OTvar{fail(}\OTvar{b}_0\OTvar{)}\OTland \OTvar{v}=\OTvar{\underline{fail}}\OTland \OTvar{b}_1=\OTvar{b}_2\\
    \OTland \OTvar{(}\OTvar{b}_1\lor\OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{\}}
    	   {\ostate{s1,t1} \xrightarrow{\OTvar{v}} \ostate{s1,t0}}
  \end{mathpar}
\caption{One of the unsatisfiable open transitions in FailureTimer}  \label{schema:unsat-ot}
\end{figure}

@@

Figure~\ref{schema:smt-lib} shows the translation of the
transition of Fig.~\ref{schema:unsat-ot}
in the SMTlib syntax.
It contains the declaration
of the BIP action algebra sorts 
and constructors, then the declaration of variables, and finally the
predicate to be checked, encoded as a set of assertions.
Here we display also the proof generated by Z3,
though in practice our tool only needs the brute information ``sat''
or ``unsat''. The result "unsat" in the output is just what we expected.
% as appears in the ``unsat'' proof generated by Z3 in which ``\{= ?x66 ?x70\}'' is shown to be false (see Fig. \ref{schema:smt-lib})

@@

\subsection{Result for the running example}
\label{section:resultOA}

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=12cm]{XFIG/FailureTimerOA}}
  \caption{Open Automaton for the Failure Monitor architecture}
  \label{schema:resultOA}
\end{figure}
